---
title: "Econometrics-4-Wei"
author: "Wei Ye^[First year Phd Student in Economics at Fordham University]"
date: "March 14,2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## a.

The code associated with a is in below:
```{r}
set.seed(1)
N=100
x=rnorm(N)
b0=1
b1=0
u=rnorm(N)
m=exp(b0+b1*x)
y=m+u
```

## b.
```{r}
theta_int=cbind(0.5,0)
source('~/Dropbox/My Mac (Wei’s MacBook Air)/Downloads/PhD-Coursework/22Spring/Econometrics/My_solution_to _hw/HW4/qfunction.R', encoding = 'UTF-8')
source('~/Dropbox/My Mac (Wei’s MacBook Air)/Downloads/PhD-Coursework/22Spring/Econometrics/My_solution_to _hw/HW4/qderivfun.R', encoding = 'UTF-8')
source('~/Dropbox/My Mac (Wei’s MacBook Air)/Downloads/PhD-Coursework/22Spring/Econometrics/My_solution_to _hw/HW4/qderivfun2.R', encoding = 'UTF-8')
out1=qderivfun(y,x,t(theta_int))
ssum=out1$ssum
out2=qderivfun2(y,x,t(theta_int))
hsum=out2$H
QgPlusOne=t(theta_int)-solve(hsum)%*%ssum
QgPlusOne

```

## c.
```{r}
library(tidyverse)
thetastore=matrix(0,nrow=2,ncol=11)
thetastore[,1]=c(0.5,0)
d=matrix(0,nrow=2,ncol=10)
for (i in 1:10){
  out1=qderivfun(y,x,thetastore[,i])
  ssum=out1$ssum
  out2=qderivfun2(y,x,thetastore[,i])
  hsum=out2$H
  thetastore[,i+1]=thetastore[,i]-solve(hsum)%*% ssum
  d[,i]=abs(thetastore[,i+1]-thetastore[,i])
}

print(thetastore)
print("\n")
print(d)
print("\n")
print(max(d))
```
From the d value matrix you can  see above, the last couple of colums' power is -12, it's extreme number, so it will converge.

Plot as below:

```{r}
plot(1:11,d[1:11])
  
```


## d:Use NLS method to estimate the model:

```{r}
nlsout=nls(y~exp(b0+b1*x),start = list(b0=1,b1=0))
summary(nlsout)
#Here the result is homoskedasticity, we dig into more to make some robust test.
```

```{r}
X1=cbind(rep(1,100),x)
o=optim(cbind(1,0),fn=qfunction,x=X1,y=y,method="BFGS")
o
```
From the result above, we can get $o\$par[,1]=2.680591$ and $o\$par[,2]=-0.001060647$, which is the same in roud 5 digits with the result we get from QgPlusOne in b.


## e. Build R script and test $\theta^{g}$ in our R script.
```{r}
##This chunk is codes for 'nrmethod.R', in this case i don't need to attach my R file anymore.
##Created by Wei Ye
##Date: March 14, 2022
#Note: I don't treat $\theta$ as vectors, but as matrix for my own convenience.
#source('~/Dropbox/My Mac (Wei’s MacBook Air)/Downloads/PhD-Coursework/22Spring/Econometrics/My_solution_to _hw/HW4/nrmethod.R', encoding = 'UTF-8')
```


Test for $\theta^{\{g\}}=[.5,0]'$

Note: I suffered from an error below in comment. I will check this tomorrow or 
in a few days, and to test it later. I have attached my nrmethod.R file in the email.

```{r}
#I comment this chunk because after implementing nrmethod to this chunk,
#there are erros: 'Only first element is used', and it follows 10 times.
#I didn't find a method to solve this at the moment. My initial suspect is due to 
#my while loop bug in "nrmethod.R".

#thetatest1=matrix(0,nrow=2,ncol=100)
#thetatest1[,1]=c(0.5,0)
#nrmethod(y,x,thetatest1)

```



```{r}
##thetatest2=matrix(0,nrow=2,ncol=100)
#thetatest2[,1]=c(-2,0)
#nrmethod(y,x,thetatest2)
```




