---
title: "Econometrics-HW5"
author: "Wei Ye"
date: "4/18/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
apple <- read_csv("apple.csv")
head(apple)


#create dummy variables
apple$ecobuy <- ifelse(apple$ecolbs>0,1,0)

attach(apple)
```
## a. Derive the log-likelihood function.

$\max l_i=p_1^{regprc_i}p_2^{ecoprc_i}(1-p_1-p_2)^{age_i}$
Take log on the above eqn:
$\max \sum_i^N \log l_i=\sum_i^N regprc_i\log p_i+ecoprc_i\log p_2+age_i\log (1-p_1-p_2)$


## b. Estimate the parameters of interest via M-estimation
```{r}
source('~/Dropbox/My Mac (Wei’s MacBook Air)/Downloads/PhD-Coursework/22Spring/Econometrics/My_solution_to _hw/HW5/qfunction.R', encoding = 'UTF-8')
 source('~/Dropbox/My Mac (Wei’s MacBook Air)/Downloads/PhD-Coursework/22Spring/Econometrics/My_solution_to _hw/HW5/qderivfun.R', encoding = 'UTF-8')
 source('~/Dropbox/My Mac (Wei’s MacBook Air)/Downloads/PhD-Coursework/22Spring/Econometrics/My_solution_to _hw/HW5/qderivfun2.R', encoding = 'UTF-8')
```

```{r}
X=cbind(regprc,ecoprc,age)
thetahat=cbind(0.2,0.6,0.2)
out <- qderivfun(ecobuy,X,t(thetahat))
out1 <- qderivfun2(ecobuy,X,t(thetahat))
out1
```


## c. Make t-test:
```{r}
ava_theta <- solve(out1$A0hat)%*% out$B0hat %*% solve(out1$A0hat)/length(ecobuy)
ava_theta
```

```{r}
se_theta <- diag(ava_theta)^(1/2)
se_theta
```
Now, make some t-tests on theta:
```{r}
t_1 <- log(thetahat)/se_theta
t_2 <- thetahat/se_theta
t_1
```
```{r}
t_2
```
No matther what, abs of t_test values are way larger than 1.96. So we can react the zero null for all variables.


## d. Use probit command in R:
```{r}
probitout <- glm(ecobuy~regprc+ecoprc+age,family=binomial(link="probit"))
probitout
```
The conclusions will not be changed.

## e. Compute the likelihood ratio test and justify whether reject the null hypothesis.
```{r}
probitout_red <- glm(ecobuy~ecoprc+age,family=binomial(link="probit"))
probitout_red
```

```{r}
#compute likelihood ratio
library(epiDisplay)
lrtest(probitout,probitout_red)
```
Actually from the command we can derive Chi-squared number from lrtest command without replying on dchisq. It's chi-squared value is high, and P-value is way lower than 0.05, therefore, we reject the null hypothesis.

