---
title: "HW6-Econometrics"
author: "Wei Ye"
date: "5/3/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(gmm)
library(tidyverse)
mauro <- read_csv("mauro1995.csv")
head(mauro)
attach(mauro)
```

## 1. State assumptions to ensure that the structural model is identified.

* $E(u)=0$, 
* $E(ux_i)=0$
* It's full rank.

## b and c. Estimations and find the moment conditions.

Source function in the files of gfunction.R, and qderivfungmm.R.
```{r}
source('~/Dropbox/My Mac (Wei’s MacBook Air)/Downloads/PhD-Coursework/22Spring/Econometrics/My_solution_to _hw/HW6/gfunction.R', encoding = 'UTF-8')
source('~/Dropbox/My Mac (Wei’s MacBook Air)/Downloads/PhD-Coursework/22Spring/Econometrics/My_solution_to _hw/HW6/qderivfungmm.R', encoding = 'UTF-8')

```

```{r}
#Below is the codes for question b and c, i met some bugs while compiling these
#code. especially in optim function. In the example code from prof. Shaw, he 
#used fn=qfunctiongmm, but we don't have this function file, and i don't know
#what's it meaning, so i changed it to fn=gfunction. However, the error is 
#Error in x%*% theta: non-conformable arguments..In order to compile this
#r-markdown file successfully, i have to comment this r block.
# N=length(gr6085)
# u=rnorm(N)
# X1=cbind(gdp60,gpop6085,sec60,corruption)
# Z1=X1
# ksi=solve((1/N)*t(Z1)%*%Z1)
# #thetah=cbind(0,0,-0.5,0.01,0)
# #Fix below later
# o = optim(cbind(0,0,0,-0.5,0),fn=gfunction,x=X1,y=corruption,z1=Z1,method="BFGS")
# b=o$par
# out1=qderivfungmm(y,x,z1,t(b),ksi)
# ksi=solve(out1$Lambda)
# o = optim(cbind(0,0,0),fn=gfunction,x=X1,y=y,z1=Z1,ksi=ksi,method="BFGS")
# b=o$par
# out1=qderivfungmm(y,x,z1,t(b),ksi)
# Avarhat=out1$Avar
# se=diag(Avarhat)^(1/2)
# t=b/se# For question c to test the null hypothesis

```



## d: Now use ptrade as an IV.

The moment condition for gmm of ptrade as an iv is that $E(ptrade_iz)=0$.

If ptrade is valide, the first state of F-stat should be larger than 10. And ptrade is not relevant to residual in the second stage.

## e: Compute the F-stat and test is it strong.

```{r}
ols_first_stage <- lm(corruption~gdp60+gpop6085+sec60+ptrade)
summary(ols_first_stage)
```
The $F-stat=2.455^2=6.03<10$, so the F-stat is not strong. And ptrade is not a good iv.

```{r}
library(ivreg)
two_stage_1 <- ivreg(gr6085~gdp60+gpop6085+sec60+corruption|gdp60+gpop6085+sec60+corruption)
summary(two_stage_1)

```

## f Use ptrade and elf as our new ivs.

```{r}
first_stage_2 <- lm(corruption~gdp60+gpop6085+sec60+elf+ptrade)
summary(first_stage_2)
```
```{r}
two_stage_2 <- ivreg(gr6085~gdp60+gpop6085+sec60+corruption|gdp60+gpop6085+sec60+ptrade+elf)
summary(two_stage_2)
```
Now, we can see based on these two methods, first on the first stage of 2sls, F-stat are still less than 10. On the other hand, the coefficient of 2sls is not statistically significant. Based on these reasons, the IVs are not strong.



