---
title: "Econoemtrics-ps3"
author: "Wei Ye"
date: "2/28/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## a:
The qfunction we modify as below:
```{r}
qfunction <- function(y,x,theta){
  N=length(y) #Find sample size
  #note that x should be N*K and theta should N*1
  m=x %*% theta#Define the function m(), %*% matrix mulplication, transpose of theta
  Q=(1/N)*sum((y-m)^2)# This is our objective function
  out=Q
}
```

## b:
Since this question requires us to derive Hessian matrix: it's just the qderivfun we coded in class, and the analytic form as below:

First,set up the q:
$$q=\frac{(y-m(x,\theta_0))^2}{2}=\frac{(y-\theta_{01}-\theta_{02}priGPA-\theta_{03}ACT)^2}{2}$$
Thus, $\frac{\partial q}{\partial \theta_{01}}=(y-m(x,\theta_0))\cdot(-1)$, $\frac{\partial q}{\partial \theta_{02}}=(\cdot)(-priGPA)$, and $\frac{\partial q}{\partial \theta_{03}}=(\cdot)(-ACT)$.
Therefore, we can derive the Hessian Matrix as:
$$H_i=\begin{bmatrix}
1&priGPA_i&ACT_i\\
priGPA_i&priGPA_i^2&priGPA_i\cdot ACT_i\\
ACT_i&ACT\cdot priGPA_i& ACT_i^2
\end{bmatrix}$$


```{r}
qderivfun<-function(y,x,thetahat){
N=length(y) #sample size
k1=dim(as.matrix(x)) #find number of x-variables
k=k1[2]+1 #add to x column to adjust for vector of ones
derivstore=matrix(0,N,k)
h=.Machine$double.eps^(1/3) #set approximation error
x=as.matrix(x) #declare as matrix
for (j in 1:k){
z=rep(0,k)
z[j]=h
for (i in 1:N){
Qph=qfunction(y[i],cbind(1,t(x[i,])),thetahat+z)
Qmh=qfunction(y[i],cbind(1,t(x[i,])),thetahat-z)
qderiv=(Qph-Qmh)/(2*h)
derivstore[i,j]=qderiv
}

}
B0hat=(1/N)*(t(derivstore)%*%derivstore) #B0hat from AVAR
ssum=colSums(derivstore)
return(list(derivstore=derivstore,ssum=ssum,B0hat=B0hat))
}
```
##c:
For computing sum of hessian:it's the file of \textbf{qderivfun2.R}, and the analytical part as below:
$$\sum_{i=1}^NH_i=\begin{bmatrix}
N&\sum_{i=1}^NpriGPA_i&\sum_{i=1}^NACT_i\\
\sum_{i=1}^NpriGPA_i&\sum_{i=1}^NpriGPA_i^2&\sum_{i=1}^NpriGPA_i\cdot ACT_i\\
\sum_{i=1}^NACT_i&\sum_{i=1}^NACT_i\cdot priGPA_i&\sum_{i=1}^N ACT_i^2
\end{bmatrix}$$
```{r}
qderivfun2 <- function(y,x,thetahat){
  k=dim(as.matrix(x))#dimension of x-variable
  p=k[2]+1 #set dimension of theta space
  N=length(y)
  derivs=vector(,N)#pre-allocate space
  H=matrix(0,p,p)#Pre-allocate space of hessian matrix
  h=.Machine$double.eps^(1/3)
  for(i in 1:p){
    for (j in 1:p) {
      z=rep(0,p)#create scalar to approximate derivative at each p
      z[j]=h
      theta1=thetahat+z
      theta2=thetahat-z
      qout1=qderivfun(y,x,theta1)
      q1=qout1$derivstor[,i]
      qout2=qderivfun(y,x,theta2)
      q2=qout2$derivstor[,i]
      d=(q1-q2)/(2*h)# Central difference
      H[i,j]=sum(d)#store the hessian matrix
    }
  }
  A0hat=(1/N)*H
  return(list(H=H,A0hat=A0hat))
}
```

##d: Compute the analytical expression  for $A_0$:
$A_0=E[H(w,\hat{\theta})]=\frac{1}{N}\sum_{i=1}^N H(w_i,\hat{\theta})$.

##e: Estimate the populaton model under consideration:
```{r}
library(tidyverse)
attend<-read_csv("attend.csv")
attach(attend)
nlsout=nls(termgpa~(b0+b1*priGPA+b2*ACT),start=list(b0=1,b1=0.04321,b2=.9))
summary(nlsout)
```
## f: Verify the result with the code qderivfun2.R:
```{r}
source('~/Dropbox/My Mac (Wei’s MacBook Air)/Downloads/PhD-Coursework/22Spring/Econometrics/My_solution_to _hw/HW3/qfunction.R', encoding = 'UTF-8')
source('~/Dropbox/My Mac (Wei’s MacBook Air)/Downloads/PhD-Coursework/22Spring/Econometrics/My_solution_to _hw/HW3/qderivfun.R', encoding = 'UTF-8')
 source('~/Dropbox/My Mac (Wei’s MacBook Air)/Downloads/PhD-Coursework/22Spring/Econometrics/My_solution_to _hw/HW3/qderivfun2.R', encoding = 'UTF-8')

thetahat=cbind(0.257753,0.875274,0.003514)
X=cbind(priGPA,ACT)
out=qderivfun(termgpa,X,t(thetahat))
out1=qderivfun2(termgpa,X,t(thetahat))
out1
```

##g: Calculate the estimated variance-covariance matrix $\hat{Avar}(\hat{\theta})$:
```{r}
avartheta=solve(out1$A0hat)%*% out$B0hat%*% solve(out1$A0hat)/length(termgpa)
avartheta
```
##h: Test the null hypothesis:
```{r}
vartheta=diag(avartheta)
se=vartheta^(1/2)
se
t_stat=thetahat/se
t_stat
```
Thus, from the results we derived above, we can't accept the alternative condition for 
$\theta_{01},\theta_{03}\neq 0$, but another $\theta$, we accept the alternative conditions that they are not equal to 0.







